Model Averaging

### 2016-06-01 model_blend.rmd
started model blending

if(.Platform$OS.type == 'windows'){
        all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
} else{
        all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
}

@@

#create test and train set: 0.75/0.25
set.seed = 100
train = sample_frac(all10mice, 0.75, replace = FALSE)
test = setdiff(all10mice, train)

#perform 10-fold cross validation on train set to create complete train set of predictions for each method
full.predict <- function(df = train, method = "log", x.folds = 5, manual.seed = 100){

        # create f.rpdol
        df$f.rpdol <- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <- "profit"
        levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <- "loss"

@@

pred.log = full.predict(df = train, method = 'log', x.folds = 5, manual.seed = 100)
pred.rf = full.predict(df = train, method = 'rf', x.folds = 5, manual.seed = 100)
pred.boost = full.predict(df = train, method = 'boost', x.folds = 5, manual.seed = 100)

@@

ggplot(gather.pred, aes(x = rownames, y = pred.val)) + geom_point(aes(colour = method))

@@

```{r}
train.blend = full_join(train %>% mutate(rown = 1:nrow(train)), pred.full, 
                        by = c('rown'='rownames'))
train.blend <- train.blend %>% select(-rown)

@@

```{r cars}
#first write formula
# create f.rpdol
train.blend$f.rpdol <- as.factor(train.blend$b.rpdol)
levels(train.blend$f.rpdol)[levels(train.blend$f.rpdol) == "0"] <- "profit"
levels(train.blend$f.rpdol)[levels(train.blend$f.rpdol) == "1"] <- "loss"

formula <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type + log + rf + boost"

#need to add interaction terms

@@

```{r pressure, echo=FALSE}
plot(pressure)
```

# Try boosted trees
```{r pressure, echo=FALSE}
plot(pressure)
```

>>>>

if(.Platform$OS.type == 'windows'){
        all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
        source("C:/Users/n9232371/Documents/github/consult/functions.R")
} else{
        all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')
}

@@

## new variables for FWLS and methods
# create f.rpdol
all10mice$f.rpdol <- as.factor(all10mice$b.rpdol)
levels(all10mice$f.rpdol)[levels(all10mice$f.rpdol) == "0"] <- "profit"
levels(all10mice$f.rpdol)[levels(all10mice$f.rpdol) == "1"] <- "loss"

#perform 10-fold cross validation on train set to create complete train set of predictions for each method
full.predict <- function(df = all10mice, method = "log", x.folds = 5, manual.seed = 100){

@@

pred.log = full.predict(df = all10mice, method = 'log', x.folds = 5, manual.seed = 100)
pred.rf = full.predict(df = all10mice, method = 'rf', x.folds = 5, manual.seed = 100)
pred.boost = full.predict(df = all10mice, method = 'boost', x.folds = 5, manual.seed = 100)

@@

ggplot(gather.pred, aes(x = rownames, y = pred.val, group = method)) + geom_point(aes(colour = method)) +
        geom_line(aes(colour = method))

@@

```{r}

all10.blend = full_join(all10mice %>% mutate(rown = 1:nrow(all10mice)), pred.full, 
                        by = c('rown'='rownames'))
all10.blend <- all10.blend %>% select(-rown)

# need to normalise method responses
ggplot(all10.blend, aes(x= sign(log) * abs(log)^(1/3))) + geom_histogram()
ggplot(all10.blend, aes(x= rf^0.5)) + geom_histogram()
ggplot(all10.blend, aes(x= log(boost))) + geom_histogram()

#cube root of log
all10.blend = mutate(all10.blend, norm.log = sign(log) * abs(log)^(1/3))
all10.blend = mutate(all10.blend, norm.rf = rf^0.5)
all10.blend = mutate(all10.blend, norm.boost = log(boost))

ggplot(all10.blend, aes(x= norm.log)) + geom_histogram()

#create test and train set: 0.75/0.25
set.seed(100)
train = sample_frac(all10.blend, 0.75, replace = FALSE)
test = setdiff(all10.blend, train)

@@

```{r cars}
#first write formula

formula.log <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type + norm.log + norm.rf + norm.boost"

formula.rf <- "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type + norm.log + norm.rf + norm.boost"

formula.boost <- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + Billing.Type + norm.log + norm.rf + norm.boost"

#formula to add interaction terms for logistic regression
metafeatures = strsplit(formula.log, split = c("\\s\\~\\s|\\s\\+\\s"))[[1]]
methods = metafeatures[13:15]
metafeatures = metafeatures[2:12]

for(j in seq_along(methods)){

        for(m in seq_along(metafeatures)){

                formula.log <- paste(formula.log, " + ", metafeatures[m], "*", methods[j], sep = "")

        }
}

#add terms to allow for constants [@Sill2009] - already done as everything is included as a single variable without interaction

#run linear regression
find.auc <- function(method = "log") {

        if (method == "log") {
                fit <- glm(as.formula(formula.log),
                               data = train,
                               family = binomial())

                pred <- predict(fit, test, type = "response")

                formula.check = "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

                check.fit <- glm(as.formula(formula.check),
                               data = train,
                               family = binomial())

                pred.c <- predict(check.fit, test, type = "response")
        }

        if (method == "rf") {
        fit <- randomForest( as.formula(formula.rf), data = train,
                mtry = 4, ntree = 1000, importance = TRUE
                )
        pred <- predict(fit, test, type = "prob")
        pred <- pred[, 2]

        formula.check = "f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"

        check.fit <- randomForest( as.formula(formula.check), data = train,
                mtry = 4, ntree = 1000
                )
        pred.c <- predict(check.fit, test, type = "prob")
        pred.c <- pred.c[, 2]
        }

        if (method == "boost") {
                fit <- gbm(as.formula(formula.boost), data = train,
                                   distribution = "bernoulli", n.trees = 6000,
                                   shrinkage = 0.0005, interaction.depth = 3,
                                   n.minobsinnode = 40)

                pred <- predict(fit, test, n.tree = 6000, type = "response")

                formula.check = "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + b.inv.log + client.totinv.log + Business + majority.pos + pc.majpos.log + Billing.Type"

                check.fit <- gbm(as.formula(formula.check), data = train,
                        distribution = "bernoulli", n.trees = 10000,
                        shrinkage = 0.001, interaction.depth = 5,
                        n.minobsinnode = 20)
                pred.c <- predict(check.fit, test, n.tree = 6000, type = "response")
        }

        #assign fit to global
        assign("fit", fit, envir = .GlobalEnv)

        #pROC
        pred.p <- pROC::roc(test$f.rpdol, pred)

        pROC::plot.roc(pred.p, col = "red", print.thres = F)

        print(pROC::auc(pred.p))

        #print check
        pred.pc <- pROC::roc(test$f.rpdol, pred.c)
        pROC::plot.roc(pred.pc, col = "blue", add = TRUE , print.thres = F)

        #print AUC's
        print(
                c("blend" = pROC::auc(pred.p), "check" = pROC::auc(pred.pc))
        )

find.auc(method = "log")
#function saves model as 'fit'
summary(fit)
#not many of the interactions had significant relationships

@@

```{r pressure, echo=FALSE}

#use caret to tune 
caret.all(df = train, formula = formula.rf, sample.frac = 0.75, seed = 100)

find.auc(method = "rf")
importance(fit, type = 1) #permutation importance
#indicates norm.log is more important

# Try boosted trees
```{r pressure, echo=FALSE}
#tune parameters

# shrinkage = 0.0005
# n.trees = 6000
# interaction.depth = 3
# n.minobsinnode = 40

find.auc(method = "boost")
summary(fit)
# norm.log was by far the most important.

Hm none of the blends are doing better than models alone.
Try comparing how similar the predictions are

```{r}

#plain model averaging

#how to normalise the prediction variables
all10.blend = mutate(all10.blend, model.av = (log + rf +boost)/3)

pred.p <- pROC::roc(all10.blend$f.rpdol, all10.blend$model.av)

pROC::plot.roc(pred.p, col = "red", print.thres = F)
pROC::plot.roc(pred.p, col = "blue", print.thres = F, add = T)

#hmm model averaging actually makes a bit of an improvement, weird.

# what about boosted tree with only 3 variables?
# or log regression wiht only 3 variables?

Try majority vote
